---
title: "Covid91 vaccine study Final2023"
author: "Marcia Sanchez"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
FinalData$treatment <- relevel(FinalData$treatment,ref="placebo")
#FinalData$infected <- relevel(FinalData$infected,ref="OK")
male_data = filter(FinalData, sex =="M")
females_data = filter(FinalData, sex =="F")
lgbtq_data = filter(FinalData, LGBTQ =="gay")
Druggies_data = filter(FinalData, DrugUser =="yes")
```



# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.


## Methods

First the finaldata variables needed to be change from character to factor variables. I used the DPLYR code for this process. Then I added the filters  to subgroups to differentiate them individually. The purpose is to see which group should take or not take vaccine.


## Males

### Graphical

```{r}
dd2 <- male_data %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```


This plot represents the male group and the comparison of the Covid921 vaccine by the vaccine takers or the placebo takers. The argument ensures that the bars for different treatments are grouped side by side for easy comparison.

The plot shows that 30% took the vaccinegot infected and 69% that received the placebo got infected .Placebo takers are a bit higher than vaccine takers.

There are two plots. In first plot,the x-axis represents infected, whether the participants were infected with covid or they were not. The y-axis represents the count, or number of people that got covid-19 or not.

In the second plot, the one above, The left two bars of the graph were all the people that didnt getcovid. The x-axis represents whether a person got covid or not. The y-axis, however, instead of representing the number of people who got covid-19, now represents the percentage of people that got covid after taking the the placebo or taking the Vaccine.


### Numerical

```{r}
table1 <- xtabs(~infected + treatment, data= male_data)
rowPerc(table1)
colPerc(table1)
```
These numerical results show specific percentages of male that got covid-19 after taking the placebo and taking the vaccine. 
Among the men, 30.43% got covid-19 after taking the vaccine, while the approximately 69.57% of men that were given the placebo contracted covid. The vaccine is clearly effective in preventing covid in males and it is recommended. Those who took the placebo were much more likely to contract covid
The odds of getting covid while taking the vaccine is 1.15: 2.55. 


### Inferential

```{r}
chisq.test(table1)
chisqtestGC(table1)
male_datafish <- fisher.test(table1)
male_datafish
```


The information to play close attention to among all of these results is the output of the Fisher's test. The p-value displayed above is extremely small at 0.00000000000001098. This is an extremely tiny number. seeing this number means that these results are not likely to occur by chance and so the null hypothesis can be rejected in favor of the alternative.
The Fishers's Test illustrates that the odds ratio is 2.24


### Effectiveness confidence Interval


```{r}
effect <- male_datafish$p.value/(male_datafish$p.value +1)
effectivenessmale = male_datafish$estimate/(male_datafish$estimate +1)*100
print(effectivenessmale)
```

```{r}
low=(male_datafish$conf.int[1]/(male_datafish$conf.int[1] +1))*100
hi=(male_datafish$conf.int[2]/(male_datafish$conf.int[2] +1))*100
options(digits=2)
```


#### Effectiveness confidence Interval `r low` to `r hi`


Above is the effectiveness for the covid-19 vaccine for male. If the effectiveness is above 50% that means that the vaccine is effective. Based on the interval seen above, we can conclude that the vaccine is ineffective in protecting males from getting covid-19. 

The odds ratio is approximately how effective the vaccine is. The odds ratio for the males can be seen above as 69%. The effectiveness confidence interval for the males is  64.26 - 73.75. It is a 95% interval.


## Females

### Graphical

```{r}
dd2 <- females_data %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```


This plot represents the female group. I am comparing Covid921 vaccine by the vaccine takers or the placebo takers. The argument ensures that the bars for different treatments are grouped side by side for easy comparison.

The second plot shows that 40% of th people that took the vaccine got infected and 60% that received the placebo got infected .Placebo takers are a bit higher than vaccine takers.

the first plot,the x-axis represents infected, whether the participants were infected with covid or they were not. The y-axis represents the count, or number of people that got covid-19 or not.

In the first plot, the one above, The left two bars of the graph were all the people that didnt getcovid at 50%. The x-axis represents whether a person got covid or not. The y-axis, however, instead of representing the number of people who got covid-19, now represents the percentage of people that got covid after taking the placebo or taking the Vaccine.


### Numerical

```{r}
table2 <- xtabs(~infected + treatment, data= females_data)
rowPerc(table2)
colPerc(table2)
```

These numerical results show specific percentages of female that got covid-19 after taking the placebo and taking the vaccine. 
Among the female, 40% got covid-19 after taking the vaccine, while the approximately 60% female were even the placebo contracted covid. The vaccine is clearly effective in preventing covid in females is recommended. Those who took the placebo were much more likely to contract covid
The odds of getting covid while taking the vaccine is 2 : 1.4

### Inferential

```{r}
chisq.test(table2)
chisqtestGC(table2)
females_datafish <- fisher.test(table2)
females_datafish
```


The information to play close attention to among all of these results is the output of the Fisher's test. The p-value displayed above is small at 0.0003. This is a small number. seeing this number means that these results are not likely to occur by chance and so the null hypothesis can be rejected in favor of the alternative.
The Fishers's Test illustrates that the odds ratio is 1.5

### Effectiveness confidence Interval


```{r}
effect <- females_datafish$p.value/(females_datafish$p.value +1)
effectivenessfemale = females_datafish$estimate/(females_datafish$estimate +1)*100
print(effectivenessfemale)
```
```{r}
low=(females_datafish$conf.int[1]/( females_datafish$conf.int[1] +1))*100
hi=(females_datafish$conf.int[2]/( females_datafish$conf.int[2] +1))*100
options(digits=2)
```


#### Effectiveness confidence Interval `r low` to `r hi`

Above is the effectiveness for the covid-19 vaccine for female. If the effectiveness is above 50% that means that the vaccine is effective. Based on the interval seen above, we can conclude that the vaccine is ineffective in protecting females from getting covid-19. 

The odds ratio is approximately how effective the vaccine is. The odds ratio for the females can be seen above as 60%. The effectiveness confidence interval for the females is 54.53 - 65.07. It is a 95% interval. I notice that female is lower than male.


## LGBTQ

### Graphical

```{r}
dd2 <- lgbtq_data %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

This plot represents the LGBTQ group. I am comparing Covid921 vaccine by the vaccine takers or the placebo takers. The argument ensures that the bars for different treatments are grouped side by side for easy comparison.

The second plot shows that 80% of the people that took the vaccine got infected and 15% that received the placebo got infected .Placebo takers are lower than vaccine takers.

The first plot,the x-axis represents infected, whether the participants were infected with covid or they were not. The y-axis represents the count, or number of people that got covid-19 or not.

In the first plot, the one above, The left two bars of the graph were all the people that didnt getcovid at 50%. The x-axis represents whether a person got covid or not. The y-axis, however, instead of representing the number of people who got covid-19, now represents the percentage of people that got covid after taking the placebo or taking the Vaccine.


### Numerical

```{r}
table3 <- xtabs(~infected + treatment, data= lgbtq_data)
rowPerc(table3)
colPerc(table3)
```

These numerical results show specific percentages of lgbtq that got covid-19 after taking the placebo and taking the vaccine. 
Among the lgbtq, 85% got covid-19 after taking the vaccine, while the approximately 15%  of lgbtq that were given the placebo contracted covid. The vaccine is not effective in preventing covid in lgbtq is not recommended. Those who took the placebo were less likely to contract covid
The odds of getting covid while taking the vaccine is 0.66 : 3.5

### Inferential

```{r}
chisq.test(table3)
chisqtestGC(table3)
lgbtq_datafish <- fisher.test(table3)
lgbtq_datafish
```


The information to play close attention to, among all of these results is the output of the Fisher's test. The p-value displayed above is extremely small at 0.00000007. This is a small number. seeing this number means that these results are not likely to occur by chance and so the null hypothesis can be rejected in favor of the alternative.
The Fishers's Test illustrates that the odds ratio is 0.18


### Effectiveness confidence Interval


```{r}
effect <- lgbtq_datafish$p.value/(lgbtq_datafish$p.value +1)
effectivenesslgbtq = lgbtq_datafish$estimate/(lgbtq_datafish$estimate +1)*100
print(effectivenesslgbtq)
```

```{r}
low=(lgbtq_datafish$conf.int[1]/( lgbtq_datafish$conf.int[1] +1))*100
hi=(lgbtq_datafish$conf.int[2]/( lgbtq_datafish$conf.int[2] +1))*100
options(digits=2)
```


#### Effectiveness confidence Interval `r low` to `r hi`

Above is the effectiveness for the covid-19 vaccine for lgbtq. If the effectiveness is above 50% that means that the vaccine is effective. Based on the interval seen above, we can conclude that the vaccine is ineffective in protecting lgbtq from getting covid-19. 

The odds ratio is approximately how effective the vaccine is. The odds ratio for the lgbtq can be seen above as 15%. The effectiveness confidence interval for the lgbtq is 26.56 to  7.60. It is a 95% interval.I notice this is lower than female and male.



## Druggies

### Graphical

```{r}
dd2 <- Druggies_data %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

This plot represents the druggies group. I am comparing Covid921 vaccine by the vaccine takers or the placebo takers. The argument ensures that the bars for different treatments are grouped side by side for easy comparison.

The second plot shows that 80% of the people that took the vaccine got infected and 14% that received the placebo got infected .Placebo takers are lower than vaccine takers.

The first plot,the x-axis represents infected, whether the participants were infected with covid or they were not. The y-axis represents the count, or number of people that got covid-19 or not.

In the first plot, the one above, The left two bars of the graph were all the people that didn't get covid at 50%. The x-axis represents whether a person got covid or not. The y-axis, however, instead of representing the number of people who got covid-19, now represents the percentage of people that got covid after taking the placebo or taking the Vaccine.


### Numerical

```{r}
table4 <- xtabs(~infected + treatment, data= Druggies_data)
rowPerc(table4)
colPerc(table4)
```

These numerical results show specific percentages of druggies that got covid-19 after taking the placebo and taking the vaccine. 
Among the druggies, 88% got covid-19 after taking the vaccine, while the approximately 12%  of druggies that were given the placebo contracted covid. The vaccine is not effective in preventing covid in druggies is not recommended. Those who took the placebo were less likely to contract covid
The odds of getting covid while taking the vaccine is 0.4 : 3.1


### Inferential

```{r}
chisq.test(table4)
chisqtestGC(table4)
Druggies_datafish <- fisher.test(table4)
Druggies_datafish
```


The information to play close attention to, among all of these results is the output of the Fisher's test. The p-value displayed above is extremely small at 0.000000003. This is an extremely small number. seeing this number means that these results are not likely to occur by chance and so the null hypothesis can be rejected in favor of the alternative.
The Fishers's Test illustrates that the odds ratio is 0.13 for druggies.


### Effectiveness confidence Interval


```{r}
effect <- Druggies_datafish$p.value/(Druggies_datafish$p.value +1)
effectivenessDruggies = Druggies_datafish$estimate/(Druggies_datafish$estimate +1)*100
print(effectivenessDruggies)
```

```{r}
low=(Druggies_datafish$conf.int[1]/( Druggies_datafish$conf.int[1] +1))*100
hi=(Druggies_datafish$conf.int[2]/( Druggies_datafish$conf.int[2] +1))*100
options(digits=2)
```



#### Effectiveness confidence Interval `r low` to `r hi`


Above is the effectiveness for the covid-19 vaccine for druggies. If the effectiveness is above 50% that means that the vaccine is effective. Based on the interval seen above, we can conclude that the vaccine is ineffective in protecting druggies from getting covid-19. 

The odds ratio is approximately how effective the vaccine is. The odds ratio for the druggies can be seen above as 11%. The effectiveness confidence interval for the druggies is 21.82 -  4.59. It is a 95% interval.It continues to decrease because is lower than male,female and lgbtq.



# Overall Results and Conclusions

The covid-19 vaccine was not effective in protecting lgbtq and druggies against the contraction of the disease. The covid-19 effectiveness was `r effectivenesslgbtq`, for lgbtq `r effectivenessDruggies`
for druggies.
The covid-19 vaccine was highly successful in protecting male and female  against covid-19.The covid-19 effectiveness was `r effectivenessmale`, `r effectivenessfemale` .Male were protected the most with this particular vaccine.
The covid-19 vaccine was not effective in treating people who identified as lgbtq. The individuals who identified as lgbtq and took the vaccine were 80%  likely to get covid-19 than those taking the placebo.The covid-19 effectiveness was `r effectivenesslgbtq` for lgbtq individuals.
The individuals that were druggies were not effectively protected by the vaccine in the same percentage as the lgbtq. 
The druggies that took the vaccine were 88% likely to get covid-19 than those taking the placebo.The covid-19 effectiveness was `r effectivenessDruggies` for druggies.

If you are a druggies or identify as lgbtq it is not recommended to receive the vaccine. However, if you are a male or a female, the vaccine was reported to have been effective.
The null that there is no difference in taking the vaccine can be accepted to the categories including lgbtq and druggies individuals. the alternative, that the vaccine works and is effective, can be accepted for the female and male.




