---
title: "Covid91 vaccine study Final2023"
author: "Marcia Sanchez"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
FinalData$treatment <- relevel(FinalData$treatment,ref="placebo")
#FinalData$infected <- relevel(FinalData$infected,ref="OK")
male_data = filter(FinalData, sex =="M")
females_data = filter(FinalData, sex =="F")
lgbtq_data = filter(FinalData, LGBTQ =="gay")
Druggies_data = filter(FinalData, DrugUser =="yes")
```




# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.


## Methods

First the finaldata variables needed to be change from character to factor variables. I used the DPLYR code for this process. Then I added the filters  to subgroups to differentiate them individually. The purpose is to see which group should take or not take vaccine.

## Males

### Graphical

```{r}
dd2 <- male_data %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```


This plot represents the male group and the comparison of the Covid921 vaccine by the vaccine takers or the placebo takers. The argument ensures that the bars for different treatments are grouped side by side for easy comparison.

The plot shows that 30% took the vaccinegot infected and 69% that received the placebo got infected .Placebo takers are a bit higher than vaccine takers.

There are two plots. In first plot,the x-axis represents infected, whether the participants were infected with covid or they were not. The y-axis represents the count, or number of people that got covid-19 or not.

In the second plot, the one above, The left two bars of the graph were all the people that didnt getcovid. The x-axis represents whether a person got covid or not. The y-axis, however, instead of representing the number of people who got covid-19, now represents the percentage of people that got covid after taking the the placebo or taking the Vaccine.


### Numerical

```{r}
table1 <- xtabs(~infected + treatment, data= male_data)
rowPerc(table1)
colPerc(table1)
```
These numerical results show specific percentages of male that got covid-19 after taking the placebo and taking the vaccine. 
Among the men, 30.43% got covid-19 after taking the vaccine, while the approximately 69.57% of men that were given the placebo contracted covid. The vaccine is clearly effective in preventing covid in males and it is recommended. Those who took the placebo were much more likely to contract covid
The odds of getting covid while taking the vaccine is 1.15: 2.55. 

### Inferential

```{r}
chisq.test(table1)
chisqtestGC(table1)
male_datafish <- fisher.test(table1)
male_datafish
```
Explain the P-value= 0.0000000000001098 There is a significance effect on the vaccine. The P- value is very low.

### Effectiveness confidence Interval


```{r}
effect <- male_datafish$p.value/(male_datafish$p.value +1)
effectivenessmale = male_datafish$estimate/(male_datafish$estimate +1)*100
print(effectivenessmale)
```

```{r}
low=(male_datafish$conf.int[1]/(male_datafish$conf.int[1] +1))*100
hi=(male_datafish$conf.int[2]/(male_datafish$conf.int[2] +1))*100
options(digits=2)
```

#### Effectiveness confidence Interval `r low` to `r hi`

The ratio is 69% and is above 50% which is effective. The confidence intervals is 64.26 - 73.75. It is a 95% interval. 




## Females

### Graphical

```{r}
dd2 <- females_data %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

This plot represents the female group. I am comparing Covid921 vaccine by the vaccine takers or the placebo takers. The argument ensures that the bars for different treatments are grouped side by side for easy comparison.

The second plot shows that 40% of th people that took the vaccine got infected and 60% that received the placebo got infected .Placebo takers are a bit higher than vaccine takers.

the first plot,the x-axis represents infected, whether the participants were infected with covid or they were not. The y-axis represents the count, or number of people that got covid-19 or not.

In the first plot, the one above, The left two bars of the graph were all the people that didnt getcovid at 50%. The x-axis represents whether a person got covid or not. The y-axis, however, instead of representing the number of people who got covid-19, now represents the percentage of people that got covid after taking the placebo or taking the Vaccine.


### Numerical

```{r}
table2 <- xtabs(~infected + treatment, data= females_data)
rowPerc(table2)
colPerc(table2)
```

These numerical results show specific percentages of female that got covid-19 after taking the placebo and taking the vaccine. 
Among the female, 40% got covid-19 after taking the vaccine, while the approximately 60% female were even the placebo contracted covid. The vaccine is clearly effective in preventing covid in females is recommended. Those who took the placebo were much more likely to contract covid
The odds of getting covid while taking the vaccine is 2 : 1.4

### Inferential

```{r}
chisq.test(table2)
chisqtestGC(table2)
females_datafish <- fisher.test(table2)
females_datafish
```

Explaining the P-value= 0.0003 There is a significance effect on the vaccine. The P- value is low.


### Effectiveness confidence Interval


```{r}
effect <- females_datafish$p.value/(females_datafish$p.value +1)
effectivenessfemale = females_datafish$estimate/(females_datafish$estimate +1)*100
print(effectivenessfemale)
```
```{r}
low=(females_datafish$conf.int[1]/( females_datafish$conf.int[1] +1))*100
hi=(females_datafish$conf.int[2]/( females_datafish$conf.int[2] +1))*100
options(digits=2)
```
#### Effectiveness confidence Interval `r low` to `r hi`

The ratio is 60% and is above 50% which is effective. The confidence intervals is 54.53 - 65.07. It is a 95% interval. I notice that famale is lower than male. 




## LGBTQ

### Graphical

```{r}
dd2 <- lgbtq_data %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

This plot represents the LGBTQ group. I am comparing Covid921 vaccine by the vaccine takers or the placebo takers. The argument ensures that the bars for different treatments are grouped side by side for easy comparison.

The second plot shows that 80% of the people that took the vaccine got infected and 15% that received the placebo got infected .Placebo takers are lower than vaccine takers.

The first plot,the x-axis represents infected, whether the participants were infected with covid or they were not. The y-axis represents the count, or number of people that got covid-19 or not.

In the first plot, the one above, The left two bars of the graph were all the people that didnt getcovid at 50%. The x-axis represents whether a person got covid or not. The y-axis, however, instead of representing the number of people who got covid-19, now represents the percentage of people that got covid after taking the placebo or taking the Vaccine.


### Numerical

```{r}
table3 <- xtabs(~infected + treatment, data= lgbtq_data)
rowPerc(table3)
colPerc(table3)
```

These numerical results show specific percentages of lgbtq that got covid-19 after taking the placebo and taking the vaccine. 
Among the lgbtq, 85% got covid-19 after taking the vaccine, while the approximately 15%  of lgbtq that were given the placebo contracted covid. The vaccine is not effective in preventing covid in lgbtq is not recommended. Those who took the placebo were less likely to contract covid
The odds of getting covid while taking the vaccine is 0.66 : 3.5

### Inferential

```{r}
chisq.test(table3)
chisqtestGC(table3)
lgbtq_datafish <- fisher.test(table3)
lgbtq_datafish
```

Explaining the P-value= 0.00000007 There is a significance effect on the vaccine. The P- value is low.


### Effectiveness confidence Interval


```{r}
effect <- lgbtq_datafish$p.value/(lgbtq_datafish$p.value +1)
effectivenesslgbtq = lgbtq_datafish$estimate/(lgbtq_datafish$estimate +1)*100
print(effectivenesslgbtq)
```

```{r}
low=(lgbtq_datafish$conf.int[1]/( lgbtq_datafish$conf.int[1] +1))*100
hi=(lgbtq_datafish$conf.int[2]/( lgbtq_datafish$conf.int[2] +1))*100
options(digits=2)
```
#### Effectiveness confidence Interval `r low` to `r hi`

The ratio is 15% and is below 50% which is not effective. The confidence intervals are 26.56 -  7.60. It is a 95% interval.


## Druggies

### Graphical

```{r}
dd2 <- Druggies_data %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

This plot represents the druggies group. I am comparing Covid921 vaccine by the vaccine takers or the placebo takers. The argument ensures that the bars for different treatments are grouped side by side for easy comparison.

The second plot shows that 80% of the people that took the vaccine got infected and 14% that received the placebo got infected .Placebo takers are lower than vaccine takers.

The first plot,the x-axis represents infected, whether the participants were infected with covid or they were not. The y-axis represents the count, or number of people that got covid-19 or not.

In the first plot, the one above, The left two bars of the graph were all the people that didn't get covid at 50%. The x-axis represents whether a person got covid or not. The y-axis, however, instead of representing the number of people who got covid-19, now represents the percentage of people that got covid after taking the placebo or taking the Vaccine.

### Numerical

```{r}
table4 <- xtabs(~infected + treatment, data= Druggies_data)
rowPerc(table4)
colPerc(table4)
```

These numerical results show specific percentages of druggies that got covid-19 after taking the placebo and taking the vaccine. 
Among the druggies, 88% got covid-19 after taking the vaccine, while the approximately 12%  of druggies that were given the placebo contracted covid. The vaccine is not effective in preventing covid in druggies is not recommended. Those who took the placebo were less likely to contract covid
The odds of getting covid while taking the vaccine is 0.4 : 3.1


### Inferential

```{r}
chisq.test(table4)
chisqtestGC(table4)
Druggies_datafish <- fisher.test(table4)
Druggies_datafish
```

Explaining the P-value= 0.000000003 There is a significance effect on the vaccine. The P- value is  very low.


### Effectiveness confidence Interval


```{r}
effect <- Druggies_datafish$p.value/(Druggies_datafish$p.value +1)
effectivenessDruggies = Druggies_datafish$estimate/(Druggies_datafish$estimate +1)*100
print(effectivenessDruggies)
```

```{r}
low=(Druggies_datafish$conf.int[1]/( Druggies_datafish$conf.int[1] +1))*100
hi=(Druggies_datafish$conf.int[2]/( Druggies_datafish$conf.int[2] +1))*100
options(digits=2)
```
#### Effectiveness confidence Interval `r low` to `r hi`


The ratio is 11% and is below 50% which is not effective. The confidence intervals are 21.82 -  4.59. It is a 95% interval.


# Overall Results and Conclusions


